---
title: "Appendix 1"
  
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: console
---
## Analysis output for "Alcohol, cannabis, and nicotine use have common and distinct associations with COVID-19 pandemic-related experiences : A Bayesian network analysis across two timepoints"

```{r, echo = F}
# formatting
library(knitr)
opts_chunk$set(tidy.opts = list(width.cutoff = 78), tidy = TRUE)
```

### This manuscript uses data from waves 9 and 16 of the COVID-Dynamic Longitudinal Study (https://coviddynamic.caltech.edu/) because those are the only waves in which both of the assessments of interest were collected (i.e., the EPII and the Monthly Substance Use Questionnaire)
### The full assessment schedule is here: https://coviddynamic.caltech.edu/documents/20311/table_W1-17_color.pdf
### The data and code are in this github repository: https://github.com/papinisan/covid-substanceuse-network

# 1. Set-up
```{r set-up}
# load packages
library(pacman)
p_load(table1, BGGM, qgraph)

# load data
df_raw <- readRDS("df_epii_mosu_916.rds")
```

# 2. Check raw data and create dataframe for analysis
```{r check-data}
# number of unique participants
length(unique(df_raw$CVDID))

# number of unique participants after removing waves in which attention checks were failed
df_raw <- df_raw[df_raw$failed == 0,]
length(unique(df_raw$CVDID))

# list of all participants
list_pts <- unique(df_raw$CVDID)

# create long dataframe that includes rows for participants who missed 1 of the 2 waves
df_pts <- data.frame(CVDID = rep(unlist(list_pts), 2), 
                     wave =rep(c(9, 16), each = length(list_pts)))
df <- merge(df_pts, df_raw, by = c("CVDID", "wave"), all.x = TRUE)

# list of nodes that will appear in the network analyses
list_nodes <- names(df[, grepl("_use|epii", names(df))])
name_nodes <- c("cannabis\nuse", "alcohol\nuse", "nicotine\nuse", "work\nimpact", 
                "home\nimpact","social\nimpact", "economic\nimpact", "emotional\nimpact", 
                "physical\nimpact", "infection\nisolation", "positive\nimpact")

# create subsets for waves 9 and 16
df9 <- df[df$wave == 9, list_nodes]
df16 <- df[df$wave == 16, list_nodes]
```

# 3. Demographic characteristics reported in the main text
1= Some high school; 2 = High school; 3 = Some college; 4 = Associate's degree; 5 = Bachelor's degree; 6 = Some graduate education; 7 = Master's degree; 8 = PhD; 9 = Professional degree
```{r demographics, warning = FALSE}
table1(~ female + age + hispanic + race + education, 
       data = df_raw[duplicated(df_raw$CVDID) == FALSE,])
```

# 4. Descriptive statistics of network variables for Table 1
```{r table-1, warning = FALSE}
table1(~ alc_use + mj_use + nic_use + epii_work + epii_home + epii_social + epii_economic + epii_emotional + epii_physical + epii_infection_isolation + epii_positive | wave, data = df)
```

# 5. Run Bayesian Graphical Gaussian Network analyses
```{r bggm}
results_compare_est <- ggm_compare_estimate(df16, df9, impute = TRUE, seed = 2022, progress = FALSE)
summary_compare_est <- summary(results_compare_est)
```

# 6. T1 (wave 9) network
```{r t1-network}
# select edges with credible intervals that do not overlap with 0
edges_t1 <- select(results_compare_est$post_samp[[2]], cred = .95)

# pcor matrix for network graph
pcor_t1 <- edges_t1$pcor_adj

# network plot
qgraph(pcor_t1, fade = T, palette = "colorblind", labels = name_nodes, negDashed = TRUE, esize = 20, cut = .05, vsize = 15, color = c(rep("#97C1A9", 3), rep("#78AED3", 8)), title = "A) July 2020")

# edge values
edges95ci_t1 <- summary(results_compare_est$post_samp[[2]], cred = .95)$dat_results
edges95ci_t1

# links between substance use and pandemic-experiences
substance_edges95ci_t1 <- edges95ci_t1[grepl("mj|nic|alc", edges95ci_t1$Relation), ]
substance_edges95ci_t1[order(substance_edges95ci_t1$Post.mean, decreasing = TRUE),]
```

# 7. T2 (wave 16) network
```{r t2-network}
# select edges with credible intervals that do not overlap with 0
edges_t2 <- select(results_compare_est$post_samp[[1]], cred = .95)

# pcor matrix for network graph
pcor_t2 <- edges_t2$pcor_adj

# network plot
qgraph(pcor_t2, fade = T, palette = "colorblind", labels = name_nodes, negDashed = TRUE, esize = 20, cut = .05, vsize = 15, color = c(rep("#97C1A9", 3), rep("#78AED3", 8)), title = "B) January 2021")

# edge values
edges95ci_t2 <- summary(results_compare_est$post_samp[[1]], cred = .95)$dat_results
edges95ci_t2

# links between substance use and pandemic-experiences
substance_edges95ci_t2 <- edges95ci_t2[grepl("mj|nic|alc", edges95ci_t2$Relation), ]
substance_edges95ci_t2[order(substance_edges95ci_t2$Post.mean, decreasing = TRUE),]
```

# 8. Export Figure 1
```{r fig1, message = FALSE}
png("figure1.png", width = 6.5, height = 3.25, units = "in", res = 300)
layout(t(1:2))
qgraph(pcor_t1, fade = T, palette = "colorblind", labels = name_nodes, negDashed = TRUE, esize = 20, cut = .05, vsize = 15, color = c(rep("#97C1A9", 3), rep("#78AED3", 8)), title = "A) July 2020")
qgraph(pcor_t2, fade = T, palette = "colorblind", labels = name_nodes, negDashed = TRUE, esize = 20, cut = .05, vsize = 15, color = c(rep("#97C1A9", 3), rep("#78AED3", 8)), title = "B) January 2021")
dev.off()
```

# 9. Compare network edges
```{r compare-networks}
summary_compare_est$dat_results
plot(summary_compare_est)
```

# 10. Environment information
```{r session-info}
sessionInfo()
```